Magnetized accretion disks around Kerr black holes with scalar hair - I. Constant angular momentum disks

### 2018-11-21 paper_prd.tex
Done. The new additions are written as comments.

\begin{table*}[t]
\caption{Central density of the disc for the different models, considering a disc with a value of $r_{\mathrm{in}}$ such as $\Delta W = 0.9 \Delta W_{\mathrm{Total}} \equiv W_{\mathrm{cusp}} - W_{\mathrm{c}}$ and a torus gravitational mass of $M_{\mathrm{T}} = 0.1 M_{\mathrm{ADM}}$. In the first column, we present the value of the central density for the different KBHsSH models. In the second column, we present the central density for a KBH with the same ADM quantities. `g.u.' stands for geometrized units.   \sg{complete caption}. \tf{Please, complete it indicating clearly what the columns report, at least the relevant ones.}}

\label{density_table}      
\centering          
\begin{tabular}{c c c c c c c c c}
\hline\hline       
 Model & $\Omega_{\mathrm{c}}$ & $\beta_{\mathrm{m_{\mathrm{c}}}}$ & $\rho_{\mathrm{c}}$ & $\rho_{\mathrm{c}}$ & $\Omega^{\rm K}_{\mathrm{c}}$ & $\rho^{\rm K}_{\mathrm{c}}$ & $\rho^{\rm K}_{\mathrm{c}}$ & $\mathfrak{s}^{\rm K}$
 \\
 & & & [g.u.] & $[\mathrm{g} \, \rm{cm}^{-3}]$ & & [g.u.] & $[\mathrm{g} \, \rm{cm}^{-3}]$ & 
  \\ 
\hline           
I & $0.493$ & $10^{10}$ & $6.818 \times 10^{-4}$ & $1.739 \times 10^{13}$ & $0.854$ & $3.752 \times 10^{-3}$ &  $9.567 \times 10^{13}$ & $1.589$ \\

 & & $1$ & $2.185 \times 10^{-3}$ & $5.503 \times 10^{13}$ & & $7.942 \times 10^{-3}$ & $2.025 \times 10^{14}$ & \\

 & & $10^{-10}$ & $3.227 \times 10^{-3}$ & $8.229 \times 10^{13}$ & & $7.641 \times 10^{-3}$ & $1.948 \times 10^{14}$ & \\

II & $0.424$ & $10^{10}$ & $3.216 \times 10^{-4}$ & $8.201 \times 10^{12}$ & $-$ & $-$ & $-$ & $-$ \\

 & & $1$ & $1.651 \times 10^{-3}$ & $4.210 \times 10^{13}$ & $-$ & $-$ & $-$ & $-$ \\

 & & $10^{-10}$ & $3.026 \times 10^{-3}$ & $7.716 \times 10^{13}$ & $-$ & $-$ & $-$ & $-$ \\

III & $0.521$ & $10^{10}$ & $8.120 \times 10^{-4}$ & $2.071 \times 10^{13}$ & $0.163$ & $6.683 \times 10^{-5}$ & $1.704 \times 10^{12}$ & $1.278$ \\

 & & $1$ & $3.497 \times 10^{-3}$ & $8.917 \times 10^{13}$ & & $2.075 \times 10^{-4}$ & $5.291 \times 10^{12}$ & \\

 & & $10^{-10}$ & $5.452 \times 10^{-3}$ & $1.390 \times 10^{14}$ & & $3.265 \times 10^{-4}$ & $8.325 \times 10^{12}$ & \\

IV & $0.427$ & $10^{10}$ & $1.197 \times 10^{-3}$ & $3.052 \times 10^{13}$ & $0.113$ & $3.001 \times 10^{-5}$ & $7.652 \times 10^{11}$ & $1.219$ \\

 & & $1$ & $3.421 \times 10^{-3}$ & $8.723 \times 10^{13}$ & & $9.512 \times 10^{-5}$ & $2.425 \times 10^{12}$ & \\

 & & $10^{-10}$ & $5.135 \times 10^{-3}$ & $1.309 \times 10^{14}$ & & $1.533 \times 10^{-4}$ & $3.909 \times 10^{12}$ & \\

V & $0.394$ & $10^{10}$ & $1.792 \times 10^{-3}$ & $4.569 \times 10^{13}$ & $0.115$ & $3.152 \times 10^{-5}$ & $8.037 \times 10^{11}$ & $1.227$\\

 & & $1$ & $3.883 \times 10^{-3}$ & $9.901 \times 10^{13}$ & & $9.942 \times 10^{-5}$ & $2.535 \times 10^{12}$ & \\

 & & $10^{-10}$ & $5.435 \times 10^{-3} $ & $1.386 \times 10^{14}$ & & $1.596 \times 10^{-4}$ & $4.070 \times 10^{12}$ & \\

VI & $0.350$ & $10^{10}$ & $2.348 \times 10^{-3}$ & $5.987 \times 10^{13}$ & $0.117$ & $3.232 \times 10^{-5}$ & $8.241 \times 10^{11}$ & $1.234$\\

 & & $1$ & $4.106 \times 10^{-3}$ & $1.047 \times 10^{14}$ & & $1.019 \times 10^{-4}$ & $2.598 \times 10^{12}$ & \\

 & & $10^{-10}$ & $5.685 \times 10^{-3}$ & $1.450 \times 10^{14}$ & & $1.632 \times 10^{-4}$ & $4.161 \times 10^{12}$ & \\

VII & $0.330$ & $10^{10}$ & $3.737 \times 10^{-3}$ & $9.529 \times 10^{13}$ & $0.129$ & $4.114 \times 10^{-5}$ & $1.049 \times 10^{12}$ & $1.268$\\

 & & $1$ & $5.356 \times 10^{-3}$ & $1.366 \times 10^{14}$ & & $1.280 \times 10^{-4}$ & $3.264 \times 10^{12}$ & \\

 & & $10^{-10}$ & $7.598 \times 10^{-3}$ & $1.937 \times 10^{14}$ & & $2.021 \times 10^{-4}$ & $5.153 \times 10^{12}$ & \\ 
 \hline\hline   
\end{tabular}
\end{table*}

@@

The dependence of the maximum specific enthalpy $h_{\mathrm{max}}$ and the maximum rest-mass density $\rho_{\mathrm{max}}$ with the magnetization parameter is shown in Fig.~\ref{comparison_HBH_Kerr_dens_enth}. The upper panels correspond to the KBHsSH models (I-VII) and the lower ones to our sequence of Kerr BHs with increasing spin parameter. For both cases, an increase in $|\Delta W|$ implies monotonically  higher values for $h_{\mathrm{max}}$ (low magnetization) and also higher values for $\rho_{\mathrm{max}}$ (high magnetization). However, there are quantitative differences between the two cases. For the enthalpy, the values of $h_{\mathrm{max}}$ reached for disks around KBHsSH are much higher than those of the Kerr BH case. This implies that, while the $w = \rho h \simeq \rho$ approximation (employed in~\cite{Komissarov:2006,Gimeno-Soler:2017}) is valid for magnetized disks ($\beta_{\mathrm{m_c}} \sim 1$) around Kerr BHs for values of the spin parameter as high as $a \sim 0.99$, that is not the case for disks around KBHsSH. We note that for the most extreme spin value we can build, $a=0.9999999$, the maximum enthalpy for the purely hydrodynamical case is $h_{\rm max}=1.692$. For this case, the maximum density in the extremely magnetized limit reaches a value of $\rho_{\mathrm{max}} = 97$, significantly larger than the value displayed in the left  panel of Fig.~\ref{comparison_HBH_Kerr_dens_enth} for the $a=0.9999$ model.

Figure~\ref{comparison_HBH_Kerr_r_m_max} shows the relative variation of the quotient of the perimeteral radius of the magnetic pressure maximum and the perimeteral radius of the disk centre, $(R_{\mathrm{m, max}}-R_{\mathrm{c}})/R_{\mathrm{c}}$, with the decimal logarithm of the magnetization parameter at the centre of the disk, $\log_{10} \beta_{\mathrm{m_c}}$. The curves plotted correspond to the same KBHsSH and Kerr BH cases as those in figure~\ref{comparison_HBH_Kerr_dens_enth}. For all cases, the radial location of the magnetic pressure maximum decreases with decreasing $\beta_{\mathrm{m_c}}$. In~\cite{Gimeno-Soler:2017} we proved that for $h=1$ disk models in stationary and axisymmetric BH spacetimes, the location of the maximum of the magnetic pressure is identical for all models when $\beta_{\mathrm{m_c}}\equiv 1 / \Gamma - 1=3$. This condition is also fulfilled for the Kerr BH case  even when $h \neq 1$, with a slight deviation for cases with very high spin parameter, \sg{Suggestion: this cannot be seen clearly in Fig.~\ref{comparison_HBH_Kerr_r_m_max}, but for spin $a = 0.9999999$, we observe $0.1912829496$ (for $a = 0.99$ is $0,01484112535$)} as can be seen in the inset of the right panel of Fig.~\ref{comparison_HBH_Kerr_r_m_max}. \tf{About the last sentence: it seems to me the condition is fulfilled completely. I do not see the deviation and if there was, could it be due to numerical precision when building the models? Have you tried to prove this analytically?} However, this condition is not fulfilled for disks built around KBHsSH (see inset in the left panel). At this point, it is relevant to remember that some of the KBHsSH models violate the Kerr bound in terms of the potential. As we mentioned previously, we need a small value of $\Delta W$ for the $h \simeq 1$ approximation to be valid in the non-magnetized regime. Now we can see that, in the KBHsSH case, this approximation is not valid even for mildly-magnetized disks.

@@

Therefore, we compute the mass of the tori around KBHsSH and, for comparison, the corresponding mass for seven disk models around Kerr BHs, each one of them with the same ADM quantities as their KBHsSH counterparts. The resulting values are reported in 
Table~\ref{density_table}. The variables corresponding to the Kerr case are indicated with a `K' superindex in this table. The fourth  and seventh columns of Table~\ref{density_table} indicate the resulting central densities for the KBHsSH and Kerr BH cases, respectively, in geometrized units. In order to compare these values with those from the end-products of binary neutron star mergers, we need to convert our results to cgs units. To this end, we first need to choose a mass for the scalar field $\mu$, as the maximum ADM mass of KBHsSH depends on $\mu$. In particular, we compute the maximum ADM mass with the following equation (see~\cite{Herdeiro:2015a} and references therein)
\begin{equation}\label{eq:max_mass_HBH}
M^{\rm max}_{\rm ADM} \simeq \alpha_{\mathrm{BS}} 10^{-19} M_{\odot} \left(\frac{\mathrm{GeV}}{\mu}\right)\,
\end{equation}
for a value of $\alpha_{\mathrm{BS}} = 1.315$ (corresponding to a value of the azimuthal harmonic index $m = 1$). The constant $\alpha_{\mathrm{BS}}$ is computed numerically for rotating boson stars. Note that Eq.~\eqref{eq:max_mass_HBH} corresponds to the maximum mass of a boson star, but, as it is mentioned in~\cite{Herdeiro:2015a}, this is also the mass for the corresponding KBHsSH.
\tf{I don't find this formula in reference~\cite{Delgado:2018}. Are you sure the formula is well written? The dimensions look strange. What is $\alpha_{\mathrm{BS}}$?}
Using a value of the mass of the scalar field of $\mu = 2.087 \times 10^{-11}$ eV yields values for the ADM mass of our models that increase from 2.043 $M_{\odot}$ to 4.799 $M_{\odot}$, from model I to VII. Note that, as is mentioned in~\cite{Ramazanoglu:2016}, the value of $\mu$ we have chose is compatible with the scalar field mass range allowed by the observational tests in Scalar-Tensor theories. \sg{not sure about this last comment, but I think something should be told about $\mu$.} \tf{Yes, but probably is not the best reference since it is not GR.}. Finally, we use the following equation
\begin{equation}
\rho_{\mathrm{cgs}} = 6.17714 \times 10^{17} \left(\frac{G}{c^2}\right)\left(\frac{M_{\odot}}{M}\right)^2 \rho_{\mathrm{geo}}\,
\end{equation}
to compute the cgs value of the central density for the different models.

>>>>

\begin{table*}[t]
\caption{Central density of the disc for the different models, considering a disc with a value of $r_{\mathrm{in}}$ such as $\Delta W = 0.9 \Delta W_{\mathrm{Total}} \equiv W_{\mathrm{cusp}} - W_{\mathrm{c}}$ and a torus gravitational mass of $M_{\mathrm{T}} = 0.1 M_{\mathrm{ADM}}$. \sg{In the third column, we present the value of the central density for the different KBHsSH models in geometrized units. In the fourth column, we present the value of the central density for the different KBHsSH models in cgs units. In the fifth and sixth columns, we present the same as in the third and fourth ones, but for a KBH with the same ADM quantities as the model displayed in the first column. The seventh column reports the sphericity of said KBH model.t. Note that `g.u.' stands for geometrized units.} . \tf{Please, complete it indicating clearly what the columns report, at least the relevant ones.}}

\label{density_table}      
\centering          
\begin{tabular}{c c c c c c c}
\hline\hline       
 Model  & $\beta_{\mathrm{m_{\mathrm{c}}}}$ & $\rho_{\mathrm{c}}$ & $\rho_{\mathrm{c}}$ & $\rho^{\rm K}_{\mathrm{c}}$ & $\rho^{\rm K}_{\mathrm{c}}$ & $\mathfrak{s}^{\rm K}$
 \\
 & & [g.u.] & $[\mathrm{g} \, \rm{cm}^{-3}]$ & [g.u.] & $[\mathrm{g} \, \rm{cm}^{-3}]$ & 
  \\ 
\hline           
I  & $10^{10}$ & $6.818 \times 10^{-4}$ & $1.739 \times 10^{13}$  & $3.752 \times 10^{-3}$ &  $9.567 \times 10^{13}$ & $1.589$ \\

  & $1$ & $2.185 \times 10^{-3}$ & $5.503 \times 10^{13}$  & $7.942 \times 10^{-3}$ & $2.025 \times 10^{14}$ & \\

  & $10^{-10}$ & $3.227 \times 10^{-3}$ & $8.229 \times 10^{13}$  & $7.641 \times 10^{-3}$ & $1.948 \times 10^{14}$ & \\

II  & $10^{10}$ & $3.216 \times 10^{-4}$ & $8.201 \times 10^{12}$ & $-$ & $-$ & $-$ \\

  & $1$ & $1.651 \times 10^{-3}$ & $4.210 \times 10^{13}$ & $-$ & $-$ & $-$ \\

  & $10^{-10}$ & $3.026 \times 10^{-3}$ & $7.716 \times 10^{13}$ & $-$ & $-$ & $-$ \\

III & $10^{10}$ & $8.120 \times 10^{-4}$ & $2.071 \times 10^{13}$  & $6.683 \times 10^{-5}$ & $1.704 \times 10^{12}$ & $1.278$ \\

  & $1$ & $3.497 \times 10^{-3}$ & $8.917 \times 10^{13}$ & $2.075 \times 10^{-4}$ & $5.291 \times 10^{12}$ & \\

  & $10^{-10}$ & $5.452 \times 10^{-3}$ & $1.390 \times 10^{14}$  & $3.265 \times 10^{-4}$ & $8.325 \times 10^{12}$ & \\

IV & $10^{10}$ & $1.197 \times 10^{-3}$ & $3.052 \times 10^{13}$  & $3.001 \times 10^{-5}$ & $7.652 \times 10^{11}$ & $1.219$ \\

  & $1$ & $3.421 \times 10^{-3}$ & $8.723 \times 10^{13}$  & $9.512 \times 10^{-5}$ & $2.425 \times 10^{12}$ & \\

  & $10^{-10}$ & $5.135 \times 10^{-3}$ & $1.309 \times 10^{14}$ & $1.533 \times 10^{-4}$ & $3.909 \times 10^{12}$ & \\

V  & $10^{10}$ & $1.792 \times 10^{-3}$ & $4.569 \times 10^{13}$  & $3.152 \times 10^{-5}$ & $8.037 \times 10^{11}$ & $1.227$\\

  & $1$ & $3.883 \times 10^{-3}$ & $9.901 \times 10^{13}$  & $9.942 \times 10^{-5}$ & $2.535 \times 10^{12}$ & \\

  & $10^{-10}$ & $5.435 \times 10^{-3} $ & $1.386 \times 10^{14}$  & $1.596 \times 10^{-4}$ & $4.070 \times 10^{12}$ & \\

VI & $10^{10}$ & $2.348 \times 10^{-3}$ & $5.987 \times 10^{13}$  & $3.232 \times 10^{-5}$ & $8.241 \times 10^{11}$ & $1.234$\\

  & $1$ & $4.106 \times 10^{-3}$ & $1.047 \times 10^{14}$  & $1.019 \times 10^{-4}$ & $2.598 \times 10^{12}$ & \\

  & $10^{-10}$ & $5.685 \times 10^{-3}$ & $1.450 \times 10^{14}$ & $1.632 \times 10^{-4}$ & $4.161 \times 10^{12}$ & \\

VII  & $10^{10}$ & $3.737 \times 10^{-3}$ & $9.529 \times 10^{13}$ & $4.114 \times 10^{-5}$ & $1.049 \times 10^{12}$ & $1.268$\\

  & $1$ & $5.356 \times 10^{-3}$ & $1.366 \times 10^{14}$ & $1.280 \times 10^{-4}$ & $3.264 \times 10^{12}$ & \\

  & $10^{-10}$ & $7.598 \times 10^{-3}$ & $1.937 \times 10^{14}$ & $2.021 \times 10^{-4}$ & $5.153 \times 10^{12}$ & \\ 
 \hline\hline   
\end{tabular}
\end{table*}

@@

The dependence of the maximum specific enthalpy $h_{\mathrm{max}}$ and the maximum rest-mass density $\rho_{\mathrm{max}}$ with the magnetization parameter is shown in Fig.~\ref{comparison_HBH_Kerr_dens_enth}. The upper panels correspond to the KBHsSH models (I-VII) and the lower ones to our sequence of Kerr BHs with increasing spin parameter. For both cases, an increase in $|\Delta W|$ implies monotonically  higher values for $h_{\mathrm{max}}$ (low magnetization) and also higher values for $\rho_{\mathrm{max}}$ (high magnetization). However, there are quantitative differences between the two cases. For the enthalpy, the values of $h_{\mathrm{max}}$ reached for disks around KBHsSH are much higher than those of the Kerr BH case. This implies that, while the $w = \rho h \simeq \rho$ approximation (employed in~\cite{Komissarov:2006,Gimeno-Soler:2017}) is valid for magnetized disks ($\beta_{\mathrm{m_c}} \sim 1$) around Kerr BHs for values of the spin parameter as high as $a \sim 0.99$, that is not the case for disks around KBHsSH. We note that for the most extreme spin value we can build, $|a-1|=10^{-7}$, the maximum enthalpy for the purely hydrodynamical case is $h_{\rm max}=1.692$. For this case, the maximum density in the extremely magnetized limit reaches a value of $\rho_{\mathrm{max}} = 97$, significantly larger than the value displayed in the left  panel of Fig.~\ref{comparison_HBH_Kerr_dens_enth} for the $a=0.9999$ model.

Figure~\ref{comparison_HBH_Kerr_r_m_max} shows the relative variation of the quotient of the perimeteral radius of the magnetic pressure maximum and the perimeteral radius of the disk centre, $(R_{\mathrm{m, max}}-R_{\mathrm{c}})/R_{\mathrm{c}}$, with the decimal logarithm of the magnetization parameter at the centre of the disk, $\log_{10} \beta_{\mathrm{m_c}}$. The curves plotted correspond to the same KBHsSH and Kerr BH cases as those in figure~\ref{comparison_HBH_Kerr_dens_enth}. For all cases, the radial location of the magnetic pressure maximum decreases with decreasing $\beta_{\mathrm{m_c}}$. In~\cite{Gimeno-Soler:2017} we proved that for $h=1$ disk models in stationary and axisymmetric BH spacetimes, the location of the maximum of the magnetic pressure is identical for all models when $\beta_{\mathrm{m_c}}\equiv 1 / \Gamma - 1=3$. This condition is also fulfilled for the Kerr BH case  even when $h \neq 1$, with a slight deviation for cases with very high spin parameter, \sg{this cannot be seen clearly in Fig.~\ref{comparison_HBH_Kerr_r_m_max}, but for spin $|a-1|=10^{-7}$, and using out ansatz radial coordinate, we observe that the deviation $r_{\mathrm{m, max}}/r_{\mathrm{c}}$, at magnetization $\beta_{\mathrm{m_{\mathrm{c}}}} = 3$, is equal to $0.9$ (in contrast, for $a = 0.9999$ its value is $0.97$).} \tf{About the last sentence: it seems to me the condition is fulfilled completely. I do not see the deviation and if there was, could it be due to numerical precision when building the models? Have you tried to prove this analytically?} However, this condition is not fulfilled for disks built around KBHsSH (see inset in the left panel). At this point, it is relevant to remember that some of the KBHsSH models violate the Kerr bound in terms of the potential. As we mentioned previously, we need a small value of $\Delta W$ for the $h \simeq 1$ approximation to be valid in the non-magnetized regime. Now we can see that, in the KBHsSH case, this approximation is not valid even for mildly-magnetized disks.

@@

Therefore, we compute the mass of the tori around KBHsSH and, for comparison, the corresponding mass for seven disk models around Kerr BHs, each one of them with the same ADM quantities as their KBHsSH counterparts. The resulting values are reported in 
Table~\ref{density_table}. The variables corresponding to the Kerr case are indicated with a `K' superindex in this table. The fourth  and seventh columns of Table~\ref{density_table} indicate the resulting central densities for the KBHsSH and Kerr BH cases, respectively, in geometrized units. In order to compare these values with those from the end-products of binary neutron star mergers, we need to convert our results to cgs units. To this end, we first need to choose a mass for the scalar field $\mu$, as the maximum ADM mass of KBHsSH depends on $\mu$. In particular, we compute the maximum ADM mass with the following equation (see~\cite{Herdeiro:2015a} and references therein)
\begin{equation}\label{eq:max_mass_HBH}
M^{\rm max}_{\rm ADM} \simeq \alpha_{\mathrm{BS}} 10^{-19} M_{\odot} \left(\frac{\mathrm{GeV}}{\mu}\right)\,
\end{equation}
\sg{for a value of $\alpha_{\mathrm{BS}} = 1.315$ (corresponding to a value of the azimuthal harmonic index $m = 1$). The constant $\alpha_{\mathrm{BS}}$ is computed numerically for rotating boson stars. Note that Eq.~\eqref{eq:max_mass_HBH} corresponds to the maximum mass of a boson star, but, as it is mentioned in~\cite{Herdeiro:2015a}, this is also the maximum mass for the corresponding KBHsSH.}
\tf{I don't find this formula in reference~\cite{Delgado:2018}. Are you sure the formula is well written? The dimensions look strange. What is $\alpha_{\mathrm{BS}}$?}
Using a value of the mass of the scalar field of $\mu = 2.087 \times 10^{-11}$ eV yields values for the ADM mass of our models that increase from 2.043 $M_{\odot}$ to 4.799 $M_{\odot}$, from model I to VII. \sg{This value of $\mu$ is within the mass range suggested by the \emph{Axiverse} of string theory (see~\cite{Arvanitaki:2010}) (a large number of scalar fields in a mass range from $10^{-33}$ eV to $10^{-10}$ eV). In addition, as is mentioned in~\cite{Ramazanoglu:2016}, the value of $\mu$ we have chose is compatible with the scalar field mass range allowed by the observational tests in Scalar-Tensor theories (Although in this paper we are working with GR, for the low values of the trace of the energy-momentum tensor $T$ we are working (in comparison with the values reached for neutron stars) both theories should be indistinguishable). We should note as well that Eq.~\cite{eq:max_mass_HBH} is valid for a non self-interacting scalar hair. Adding self-interaction terms will produce astrophysically relevant solutions for less exotic values of the scalar field mass $\mu$ (see~\cite{Delgado:2018}).} \tf{Yes, but probably is not the best reference since it is not GR.} \sg{Once we have computed the new values of the central density, we use the following equation (from~\cite{RezzollaBook})}
\begin{equation}
\rho_{\mathrm{cgs}} = 6.17714 \times 10^{17} \left(\frac{G}{c^2}\right)\left(\frac{M_{\odot}}{M}\right)^2 \rho_{\mathrm{geo}}\,
\end{equation}
\sg{to obtain the value of the central density in cgs units for the different models.}
